Numerical Stability of Detonations in White Dwarf Simulations

### 2018-07-10 papers/ignition/paper.tex
Start reworking the introduction

Binary white dwarf (WD) systems are a promising progenitor candidate for Type
Ia supernovae (SNe Ia) for observational reasons, but from the simulation
perspective there is still significant uncertainty regarding how nuclear
detonations form in an exploding white dwarf. There is uncertainty in the
model, regarding whether detonations form on the surface of a star as
accreting material heats up when it strikes the surface, or at or near the
center of the star as the accreted material compresses the WD. There is also
uncertainty in the numerics: given one of these models, can we self-consistently
demonstrate that a detonation occurs? This question has haunted both single
WD and double WD progenitor models, because the length and time scale at which
a detonation forms is orders of magnitude smaller than the resolution that
typical hydrodynamic simulations can achieve. Furthemore, the approximate equations
we use at the resolution that we can achieve are still notoriously
difficult to solve accurately. The mere presence of a detonation
(or lack thereof) in a simulation is therefore only weak evidence regarding
whether a detonation would truly occur.

>>>>

Thermonuclear detonations are common to all current likely models of Type Ia
supernovae (SNe Ia), but how they are actually generated in progenitor systems
is still an open question. Different models predict different locations for
the detonation (e.g. in the center of a star or on the edge) and different
mechanisms for initiating the event (e.g. a prompt detonation in a white dwarf
collision versus a deflagration to detonation transition in the single degenerate
model). Common to all of the cases is a severe lack of numerical resolution in
the location where the detonation is expected to occur. The length and time scale
at which a detonation forms is orders of magnitude smaller than the resolution that
typical hydrodynamic simulations can achieve. Furthemore, the approximate equations
we use at the resolution that we can achieve are still notoriously difficult to
solve accurately. The mere presence of a detonation (or lack thereof) in a simulation
is therefore only weak evidence regarding whether a detonation would truly occur.
